---
title: "Exploring nutrients or environmental indicators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring nutrients or environmental indicators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  results = "hide"
)
```

# Load libraries
```{r setup, message = FALSE}
library(sustainableNutRients)
library(tidyverse)
library(ggbeeswarm)
library(gghalves)

theme_set(theme_bw())
```

## Calculate nutritional content and environmental content per portion
The data used here are the recipes from Oda used in "Map foods to database".
```{r read-data, echo=FALSE, message=FALSE, warning=FALSE}
# Read the data
data("recipes_oda")
data(composite_ingredients_oda)

#Additional entries to databases
additional_entries <- list(
  "volume_weight" = tibble(
    tmp = c("bolognese_vegetarian pcs 370",
            "hummus pcs 165",
            "pizza_dough pcs 210",
            "sauce_pizza pack 400",
            "sauce_pad pcs 150",
            "soup_tomato pcs 530",
            "vanilla_pod pcs 3",
            "fish balls_curry sauce pcs 800")
  ) %>% separate(col = tmp, into = c("recipe_name", "unit", "grams_per_unit"), sep = " ") %>%
    mutate(grams_per_unit = as.numeric(grams_per_unit)) %>%
    createNewDatabaseEntry(., "volume_weight"),

  "nutrients" = composite_ingredients_oda %>%
    #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
   standardiseFoodList() %>%
    findFoodInDatabase(., "volume_weight") %>%
    calculateWeightOfIngredients() %>%
    findFoodInDatabase(., "nutrients") %>%
    #Calculate nutrients per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
    calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g") %>%
    #Format as new database entries for the nutrients database
    createNewDatabaseEntry(., "nutrients"),

  "environmental_impact" = composite_ingredients_oda %>%
   #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
    standardiseFoodList() %>%
    findFoodInDatabase(., "volume_weight") %>%
    calculateWeightOfIngredients() %>%
    findFoodInDatabase(., "sustainability") %>%
    #Calculate env impact per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
    calculateEnvironmentalImpactOfFoodlist(., calculate_sustainability = "per 100 g") %>%
    #Format as new database entries for the nutrients database
    createNewDatabaseEntry(., "sustainability")
)

#Run twice since some composite ingredients are used in other composite ingredients
additional_entries$nutrients <-  composite_ingredients_oda %>%
    #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
    standardiseFoodList() %>%
    findFoodInDatabase(., "volume_weight") %>%
    calculateWeightOfIngredients() %>%
    findFoodInDatabase(., "nutrients", additional_entries = additional_entries$nutrients) %>%
    #Calculate nutrients per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
    calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g") %>%
    #Format as new database entries for the nutrients database
    createNewDatabaseEntry(., "nutrients")

additional_entries$environmental_impact <- composite_ingredients_oda %>%
   #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
    standardiseFoodList() %>%
    findFoodInDatabase(., "volume_weight") %>%
    calculateWeightOfIngredients() %>%
    #Calculate env impact per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
    #Run twice since some of the ingredients are used in the composite ingredients themselves
    findFoodInDatabase(., "sustainability", additional_entries = additional_entries$environmental_impact) %>%
    calculateEnvironmentalImpactOfFoodlist(., calculate_sustainability = "per 100 g") %>%
    #Format as new database entries for the nutrients database
    createNewDatabaseEntry(., "sustainability")
```

```{r calculate-ingredient-values}
#Calculate nutrient content and environmental impact per portion
#First standardise recipe ingredient names and amounts and map recipe ingredients to databases
standardised_recipe <- recipes_oda$recipes %>%
    mutate(Ingredients = Ingredients %>%
             str_replace("slider bread", "hamburger bread") %>%
             str_replace(" stalk ", "pcs")) %>%
    standardiseFoodList()
mapped_to_database <- list(
  "nutrients" = standardised_recipe %>%
    #Fix name of an ingredient
    mutate(Ingredients = case_when(
      Original_ingredients == "fish balls in curry sauce ready-made" ~ "fish balls in curry sauce",
      TRUE ~ Ingredients)) %>%
    #Map to volume_weight database
    findFoodInDatabase(., "volume_weight", additional_entries = additional_entries$volume_weight) %>%
    calculateWeightOfIngredients() %>%
    #Map tp nutrient database,add number of portions and calculate nutrient content per portion
    findFoodInDatabase(., "nutrients", additional_entries = additional_entries$nutrients) %>%
    mutate(database_ID = case_when(
      Ingredients == "fish balls in curry sauce" ~ fixFoodMappingError(additional_entries$nutrients$query_words, "fish ball", "curry sauce"),
      Ingredients == "sausage stew" ~ fixFoodMappingError(additional_entries$nutrients$query_words, "sausage", "stew"),
      Ingredients == "stew mix" ~ fixFoodMappingError(additional_entries$nutrients$query_words, "stew", "mix"),
      TRUE ~ database_ID)),
  
  "environmental_impact" = standardised_recipe %>%
    mutate(Ingredients = case_when(
    Original_ingredients == "fish balls in curry sauce ready-made" ~ "fish balls in curry sauce",
    TRUE ~ Ingredients)) %>%
  #Map to volume_weight database
  findFoodInDatabase(., "volume_weight", additional_entries = additional_entries$volume_weight) %>%
  calculateWeightOfIngredients() %>%
  #Map to SHARP ID, add number of portions and calculate environmental impact per portion
  findFoodInDatabase(., "sustainability", additional_entries = additional_entries$environmental_impact)  %>%
  mutate(database_ID = case_when(
    Ingredients == "fish balls in curry sauce" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "fish ball", "curry sauce"),
    Ingredients == "sausage stew" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "sausage", "stew"),
    Ingredients == "stew mix" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "stew", "mix"),
    str_detect(Ingredients, "mac and cheese") ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "mac and cheese", "bacon"),
    Ingredients == "bean salad" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "bean salad"),
    Ingredients == "bean taco" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "bean", "taco"),
    Ingredients == "chili sin carne ready made" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "chili", "sin carne"),
    Ingredients == "meatballs in tomato sauce" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "meatball", "tomato sauce"),
    TRUE ~ database_ID))
)
#Calculate nutrient content and environmental impact per portion
calculations <- list()
calculations$nutrient_content <- mapped_to_database$nutrients %>%
  #Add number of portions
  left_join(., recipes_oda$meta %>% select(recipe_name, number_of_portions) %>% unique()) %>%
  calculateNutrientContentOfFoodlist(., calculate_nutrients = "per portion")

calculations$environmental_impact <- mapped_to_database$environmental_impact %>%
  #Add number of portions
  left_join(., recipes_oda$meta %>% select(recipe_name, number_of_portions) %>% unique()) %>%
  calculateEnvironmentalImpactOfFoodlist(., calculate_sustainability = "per portion")
```

```{r plots-protein-sources, include = FALSE}
# 
# protein_sources <- recipes_oda$recipes %>%
#   standardiseFoodList() %>%
#   #Map to SHARP ID, add number of portions and calculate environmental impact per portion
#   findFoodInDatabase(., "sustainability", additional_entries = additional_entries$environmental_impact) %>%
#   #Add foodgroups
#   left_join(., sustainableNutRients:::SHARP2018_foodgroups) %>%
#   #Find protein source
#   select(Ingredients, foodgroup, recipe_name) %>% unique() %>%
#   #Separate into beef, lamb, pork, poultry, lean fish, oily fish, shellfish, vegetarian, vegan
#   #Animal sourced foods
#   mutate(protein = case_when(
# 
#     str_detect(Ingredients, 'beef|meatball') & !str_detect(Ingredients, 'fund|broth|gravy|elk|deer|venison') ~ 'beef',
#     str_detect(Ingredients, 'reindeer|elk|rabbit|grouse|roe deer|venison') ~ 'game',
#     str_detect(Ingredients, 'lamb') & !str_detect(Ingredients, 'broth|salad') ~ 'lamb',
#     str_detect(Ingredients, 'pork|bacon|sausage|ham|salami') & !str_detect(Ingredients, 'lard|beef|burger') ~ 'pork',
#     str_detect(Ingredients, 'chicken|turkey|duck|goose') & !str_detect(Ingredients, 'broth|fat|sauce') ~ 'poultry',
# 
#     str_detect(Ingredients, 'pollock|cod|anglerfish|fish cakes coarse|fish ball|fish stick|haddock|grouper|catfish|sea bass|halibut|tuna') ~ 'lean fish', #The fishcakes are made out of mostly haddock
#     str_detect(Ingredients, 'salmon|trout|arctic char|mackerel|herring|sardine|anchovy') & !str_detect(Ingredients, 'roe') ~ 'oily fish',
#     str_detect(Ingredients, 'squid|prawn|shrimp|mussel|crab|lobster|shellfish|scampi|clam|scallop') & !str_detect(Ingredients, 'paste') ~ 'shellfish'
# 
#     )) %>%
#   #Vegetarian or vegan foods
#   group_by(recipe_name, foodgroup) %>%
#   mutate(
#     protein = case_when(
# 
#       !any(foodgroup %in% c('Seafood', 'Eggs', 'Milk and dairy products', 'Meat and nmeat products', 'Fish, seafood, amphibians, reptiles and invertebrates')) &
#         !str_detect(Ingredients, 'lamb|game|fish|beef|chicken|turkey|shrimp|cheese|condensed cream|egg\\b|duck|honey|pork|mayonnaise|barbeque|oyster|worcestershire|sausage|ham\\b|shortening|sour cream|tzatziki') &
#         !Ingredients %in% c('butter', 'unsalted butter', 'butter clarified ghee', 'butter for cooking',
#                             'buttermilk', 'refrigerated buttermilk biscuit dough', 'spice butter', 'puff pastry',
#                             'shop-bought shortcrust pastry') ~ 'plants',
#       !any(foodgroup %in% c('Seafood', 'Meat and\nmeat products')) & !str_detect(Ingredients, 'condensed cream|beef|fish|shrimp|duck|sausage|shortening') ~ 'vegetarian',
#       TRUE ~ protein),
#     vegetarian = case_when(
#       protein %in% c('plants', 'vegetarian') ~ 'vegetarian',
#       TRUE ~ protein
# 
#       )) %>% group_by(recipe_name) #%>%
#     #Clean up
#     # summarise(type = case_when(
#     #   any(protein == 'beef') ~ 'Beef',
#     #   any(protein == 'game') ~ 'Game',
#     #   any(protein == 'lamb') ~ 'Lamb',
#     #   any(protein == 'poultry') ~ 'Poultry',
#     #   any(protein == 'lean fish') ~ 'Lean fish',
#     #   any(protein == 'oily fish') ~ 'Oily fish',
#     #   any(protein == 'shellfish') ~ 'Shellfish',
#     #   any(protein == 'pork') ~ 'Pork',
#     #   all(protein == 'plants') ~ 'Vegan',
#     #   all(vegetarian == 'vegetarian') ~ 'Vegetarian'
#     #   )) %>% ungroup()

```

## Plot data
### Nutrient content
Looking at how much of the recommended intake of nutrient found per portion of the recipes.
```{r nutrients-rdi, fig.width=8, fig.height=5}
# Recommended intake of nutrients for adult women in Norway
nutrient_units <- tibble(
  'nutrient' = c('Vitamin A', 'Retinol', 'Beta-carotene',
                'Vitamin D','Vitamin E', 'Thiamin',
                'Riboflavin','Niacin', 'Vitamin B6',
                'Folate', 'Vitamin B12', 'Vitamin C',
                'Calcium', 'Iron', 'Sodium',
                'Potassium', 'Magnesium', 'Zinc',
                'Selenium', 'Copper', 'Phosphorus',
                'Iodine'),  
  'rdi' = c(700, 700, 8400,
            10, 8, 1.1,
            1.3, 15, 1.2,
            400, 2, 75,
            800, 15, 2300,
            3100, 280, 7,
            50, 0.9, 600,
            150))

#Plot function
buildPlot <- function(dataframe, title, x_axis, x_axis_title, y_axis, y_axis_title, ylim, title_text_size = 12) {
  ggplot(dataframe, aes(x = .data[[x_axis]], y = .data[[y_axis]])) +
    geom_half_boxplot(side = "r", outlier.shape = NA) +
    geom_half_violin() +
    geom_quasirandom(alpha = 0.5, size = 2, shape = 21) +
    labs(
      title = title,
      x = x_axis_title,
      y = y_axis_title
    ) +
    coord_cartesian(ylim = ylim) +
    theme(plot.title = element_text(size = title_text_size))
}

# Calculate pct of intake per portion
rdi_per_portion <- calculations$nutrient_content %>%
  right_join(nutrient_units) %>%
  mutate(pct_rdi_per_portion = (nutrient_content_per_portion/rdi)*100)

#Minerals
buildPlot(dataframe = rdi_per_portion %>% filter(nutrient %in% c(
  'Calcium', 'Iron', 'Sodium',
  'Potassium', 'Magnesium', 'Zinc',
  'Selenium', 'Copper', 'Phosphorus',
  'Iodine')), 
  title = "Percent of recommended daily intake of minerals found per portion of the recipes", x_axis = "nutrient", x_axis_title = "Mineral", y_axis = "pct_rdi_per_portion", y_axis_title = "Percent of RDI", ylim = c(0, 600)) #Cut out a few outliers to make plot more readable
# Water soluble vitamins
buildPlot(dataframe = rdi_per_portion %>% filter(nutrient %in% c(
  'Thiamin', 'Riboflavin','Niacin', 'Vitamin B6',
  'Folate', 'Vitamin B12', 'Vitamin C')), 
  title = "Percent of recommended daily intake of water soluble vitamins found per portion of the recipes", x_axis = "nutrient", x_axis_title = "Water soluble vitamin", y_axis = "pct_rdi_per_portion", y_axis_title = "Percent of RDI", ylim = c(0, 500))
# Fat soluble vitamins
buildPlot(dataframe = rdi_per_portion %>% filter(nutrient %in% c(
  'Vitamin A', 'Retinol', 'Beta-carotene', 'Vitamin D','Vitamin E')), 
  title = "Percent of recommended daily intake of fat soluble vitamins found per portion of the recipes", x_axis = "nutrient", x_axis_title = "Fat soluble vitamin", y_axis = "pct_rdi_per_portion", y_axis_title = "Percent of RDI", ylim = c(0, 750))
```
Some of these values are very high, likely due to some recipes being said to be only one portion, while they are actually meant for more people. These are typically recipes for baked goods such as waffles, cookies etc.

### Environmental impact of the recipes
Using ggplot to visualise the environmental impact of the recipes 
```{r nutrient-content-env-impact_per_portion, fig.width=5, fig.height=5}
#Environmental impact per portion
buildPlot(dataframe = calculations$environmental_impact, 
  title = "Environmental impact per portion of Oda recipes SHARP ID", x_axis = "environmental_impact_indicator", x_axis_title = "Environmental impact indicator", y_axis = "environmental_impact_per_portion", y_axis_title = "Environmental impact in", ylim = c(0, 25), title_text_size = 11) +
  facet_wrap(~environmental_impact_indicator, scales = "free") 
```


Similarly with the nutrient content, some of these values are very high. Looking only are recipes with more than one portion gives a slightly different view:
```{r nutrient-content-env-impact_per_portion2, fig.width=5, fig.height=5}
# Recipes with only one portion have a larger environmental impact than recipes with more than 1 portion, look only at those.
buildPlot(dataframe = calculations$environmental_impact %>% filter(number_of_portions != 1), 
  title = "Environmental impact per portion of Oda recipes \nwith number of portions >1 mapped to SHARP ID", x_axis = "environmental_impact_indicator", x_axis_title = "Environmental impact indicator", y_axis = "environmental_impact_per_portion", y_axis_title = "Environmental impact", ylim = c(0, 7), title_text_size = 11) +
  facet_wrap(~environmental_impact_indicator, scales = "free") 
```

### Nutriscore
The Nutriscore is a front of pack label that should give consumers at-a-glance information about a foods healthiness. The Nutriscore is calculated per 100 g, from the amount of energy, macronutrients and percentage of fruit, vegetables, legumes, nuts and healthy oils in a food. The necessary information can be gained from calculating the nutrient content per 100 g of a recipe, and then using the food groups from SHARP to find ingredients that are fruit, vegetables or legumes and individual ingredient names to identify the healthy oils.
```{r nutriscore-dataframe, include = FALSE}
calculations$nutrient_content_100g <- mapped_to_database$nutrients %>%
  calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g")
```
```{r nutriscore-calculation}
# Calculate percentage of fruit/vegetables/legumes/nutes/oil in a recipe
# Weight of all ingredients, use ingredients mapped to SHARP db to use the foodgroups in SHARP later
tmp <- mapped_to_database$environmental_impact %>%
  #Only look at recipes that have no composite ingredients for simplicity's sake
  group_by(recipe_name) %>% 
  mutate(to_remove = case_when(any(str_detect(database_ID, ".999")) ~ "remove", TRUE ~ "keep")) %>%
  ungroup() %>%
  filter(to_remove == "keep") %>% select(-to_remove) %>%
  #Use foodgroups from SHARP to identify fruit, vegetables and legumes
  left_join(., SHARP2018_foodgroups) %>%
  select(recipe_name, Ingredients, Amounts, foodgroup) %>% distinct() %>%
  # #Mark the foods relevant for Nutriscore
  mutate(
     nutriscore_foods = case_when(
       (foodgroup %in% c('Vegetables and vegetable products', 'Fruit and fruit products', 'Fruit/vegetable juice and nectar', 'Legumes, nuts, oilseeds and spices') & !str_detect(Ingredients, 'mushroom|coconut milk|sesame seed|pine')) | str_detect(Ingredients, 'olive oil|rapeseed oil|walnut oil') ~ 'nutriscore_fruit_veg_legumes_nuts_oils',
       TRUE ~ "Other")
    ) %>%
  # Calculate percent of total recipe weight
  #Amounts of dried fruit/veg and ketchup/purees should be doubled as per Nutriscore guidelines
  mutate(
      Amounts = case_when(
        nutriscore_foods == "nutriscore_fruit_veg_legumes_nuts_oils" & str_detect(Ingredients, 'paste tomato|ketchup|dried|raisin|prune') ~ Amounts*2,
        TRUE ~ Amounts)) %>%
  group_by(recipe_name, nutriscore_foods) %>%
  mutate(weight_foodgroups = sum(Amounts, na.rm = TRUE)) %>% ungroup() %>%
  select(recipe_name, weight_foodgroups, nutriscore_foods) %>% distinct() %>%
  group_by(recipe_name) %>%
  #Use same column and variable names as nutrient content dataframe
  mutate(
    nutrient_content_per_hektogram = (weight_foodgroups/sum(weight_foodgroups, na.rm = TRUE))*100,
         nutrient = "nutriscore_amounts_pct") %>% ungroup() %>%
  filter(nutriscore_foods != "Other") %>%
  select(recipe_name, nutrient, nutrient_content_per_hektogram)

#Add nutrients and calculate Nutriscore
nutriscore <- tmp %>%
  #Only add nutrients from recipes with composite ingredients
  bind_rows(., calculations$nutrient_content_100g %>%
              filter(recipe_name %in% tmp$recipe_name)
            ) %>%
  #Calculate nutriscore
  calculateNutriscore()
```
Plotting the results
```{r nutriscore-plots, fig.width=6, fig.height=4}
ggplot(nutriscore, aes(x = nutriscore_letter, fill = nutriscore_letter)) +
  geom_bar(show.legend = FALSE) +
  #Add numbers above bars
  geom_text(stat='count', aes(label=after_stat(count)), size = 3, vjust=-1) +
  #Increase y axis to allow for text
  coord_cartesian(ylim = c(0, 15)) +
  #Add titles
  labs(
    title = "Nutriscore of a subset of Oda recipes",
    y = "Number of recipes",
    x = "Nutriscore letter"
  ) +
  #Use Nutriscore coloring scheme
  scale_fill_manual(
    values = c("A" = "#028141", "B" = "#85BB2F", "C" = "#FECB03",
               "D" = "#EE8100", "E" = "#E63D11"))

```
