---
title: "Exploring nutrients or environmental indicators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exploring nutrients or environmental indicators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  results = "hide"
)
```

# Load libraries
```{r setup, message = FALSE}
library(sustainableNutRients)
library(tidyverse)
library(ggbeeswarm)
library(gghalves)

theme_set(theme_bw())
```

## Calculate nutritional content and environmental content per portion
The data used here are the recipes from Oda used in "Map foods to database".
```{r read-data, echo=FALSE, message=FALSE, warning=FALSE}
# Read the data
data("recipes_oda")
data(composite_ingredients_oda)

#Additional entries to databases
additional_entries <- list(
  "volume_weight" = tibble(
    tmp = c("bolognese_vegetarian pcs 370",
            "hummus pcs 165",
            "pizza_dough pcs 210",
            "sauce_pizza pack 400",
            "sauce_pad pcs 150",
            "soup_tomato pcs 530",
            "vanilla_pod pcs 3",
            "fish balls_curry sauce pcs 800")
  ) %>% separate(col = tmp, into = c("recipe_name", "unit", "grams_per_unit"), sep = " ") %>%
    mutate(grams_per_unit = as.numeric(grams_per_unit)) %>%
    createNewDatabaseEntry(., "volume_weight"),

  "nutrients" = composite_ingredients_oda %>%
    #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
   standardiseFoodList() %>%
    findFoodInDatabase(., "volume_weight") %>%
    calculateWeightOfIngredients() %>%
    findFoodInDatabase(., "nutrients") %>%
    #Calculate nutrients per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
    calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g") %>%
    #Format as new database entries for the nutrients database
    createNewDatabaseEntry(., "nutrients"),

  "environmental_impact" = composite_ingredients_oda %>%
   #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
    standardiseFoodList() %>%
    findFoodInDatabase(., "volume_weight") %>%
    calculateWeightOfIngredients() %>%
    findFoodInDatabase(., "sustainability") %>%
    #Calculate env impact per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
    calculateEnvironmentalImpactOfFoodlist(., calculate_sustainability = "per 100 g") %>%
    #Format as new database entries for the nutrients database
    createNewDatabaseEntry(., "sustainability")
)

#Run twice since some composite ingredients are used in other composite ingredients
additional_entries$nutrients <-  composite_ingredients_oda %>%
    #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
    standardiseFoodList() %>%
    findFoodInDatabase(., "volume_weight") %>%
    calculateWeightOfIngredients() %>%
    findFoodInDatabase(., "nutrients", additional_entries = additional_entries$nutrients) %>%
    #Calculate nutrients per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
    calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g") %>%
    #Format as new database entries for the nutrients database
    createNewDatabaseEntry(., "nutrients")

additional_entries$environmental_impact <- composite_ingredients_oda %>%
   #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
    standardiseFoodList() %>%
    findFoodInDatabase(., "volume_weight") %>%
    calculateWeightOfIngredients() %>%
    #Calculate env impact per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
    #Run twice since some of the ingredients are used in the composite ingredients themselves
    findFoodInDatabase(., "sustainability", additional_entries = additional_entries$environmental_impact) %>%
    calculateEnvironmentalImpactOfFoodlist(., calculate_sustainability = "per 100 g") %>%
    #Format as new database entries for the nutrients database
    createNewDatabaseEntry(., "sustainability")
```

```{r calculate-ingredient-values}
#Calculate nutrient content and environmental impact per portion
#First standardise recipe ingredient names and amounts and map recipe ingredients to databases
standardised_recipe <- recipes_oda$recipes %>%
    mutate(Ingredients = Ingredients %>%
             str_replace("slider bread", "hamburger bread") %>%
             str_replace(" stalk ", "pcs")) %>%
    standardiseFoodList()
mapped_to_database <- list(
  "nutrients" = standardised_recipe %>%
    #Fix name of an ingredient
    mutate(Ingredients = case_when(
      Original_ingredients == "fish balls in curry sauce ready-made" ~ "fish balls in curry sauce",
      TRUE ~ Ingredients)) %>%
    #Map to volume_weight database
    findFoodInDatabase(., "volume_weight", additional_entries = additional_entries$volume_weight) %>%
    calculateWeightOfIngredients() %>%
    #Map tp nutrient database,add number of portions and calculate nutrient content per portion
    findFoodInDatabase(., "nutrients", additional_entries = additional_entries$nutrients) %>%
    mutate(database_ID = case_when(
      Ingredients == "fish balls in curry sauce" ~ fixFoodMappingError(additional_entries$nutrients$query_words, "fish ball", "curry sauce"),
      Ingredients == "sausage stew" ~ fixFoodMappingError(additional_entries$nutrients$query_words, "sausage", "stew"),
      Ingredients == "stew mix" ~ fixFoodMappingError(additional_entries$nutrients$query_words, "stew", "mix"),
      TRUE ~ database_ID)),
  
  "environmental_impact" = standardised_recipe %>%
    mutate(Ingredients = case_when(
    Original_ingredients == "fish balls in curry sauce ready-made" ~ "fish balls in curry sauce",
    TRUE ~ Ingredients)) %>%
  #Map to volume_weight database
  findFoodInDatabase(., "volume_weight", additional_entries = additional_entries$volume_weight) %>%
  calculateWeightOfIngredients() %>%
  #Map to SHARP ID, add number of portions and calculate environmental impact per portion
  findFoodInDatabase(., "sustainability", additional_entries = additional_entries$environmental_impact)  %>%
  mutate(database_ID = case_when(
    Ingredients == "fish balls in curry sauce" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "fish ball", "curry sauce"),
    Ingredients == "sausage stew" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "sausage", "stew"),
    Ingredients == "stew mix" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "stew", "mix"),
    str_detect(Ingredients, "mac and cheese") ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "mac and cheese", "bacon"),
    Ingredients == "bean salad" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "bean salad"),
    Ingredients == "bean taco" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "bean", "taco"),
    Ingredients == "chili sin carne ready made" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "chili", "sin carne"),
    Ingredients == "meatballs in tomato sauce" ~ fixFoodMappingError(additional_entries$environmental_impact$query_words, "meatball", "tomato sauce"),
    TRUE ~ database_ID))
)
#Calculate nutrient content and environmental impact per portion
calculations <- list()
calculations$nutrient_content <- mapped_to_database$nutrients %>%
  #Add number of portions
  left_join(., recipes_oda$meta %>% select(recipe_name, number_of_portions) %>% unique()) %>%
  calculateNutrientContentOfFoodlist(., calculate_nutrients = "per portion")

calculations$environmental_impact <- mapped_to_database$environmental_impact %>%
  #Add number of portions
  left_join(., recipes_oda$meta %>% select(recipe_name, number_of_portions) %>% unique()) %>%
  calculateEnvironmentalImpactOfFoodlist(., calculate_sustainability = "per portion")
```

```{r plots-protein-sources, include = FALSE}
# 
# protein_sources <- recipes_oda$recipes %>%
#   standardiseFoodList() %>%
#   #Map to SHARP ID, add number of portions and calculate environmental impact per portion
#   findFoodInDatabase(., "sustainability", additional_entries = additional_entries$environmental_impact) %>%
#   #Add foodgroups
#   left_join(., sustainableNutRients:::SHARP2018_foodgroups) %>%
#   #Find protein source
#   select(Ingredients, foodgroup, recipe_name) %>% unique() %>%
#   #Separate into beef, lamb, pork, poultry, lean fish, oily fish, shellfish, vegetarian, vegan
#   #Animal sourced foods
#   mutate(protein = case_when(
# 
#     str_detect(Ingredients, 'beef|meatball') & !str_detect(Ingredients, 'fund|broth|gravy|elk|deer|venison') ~ 'beef',
#     str_detect(Ingredients, 'reindeer|elk|rabbit|grouse|roe deer|venison') ~ 'game',
#     str_detect(Ingredients, 'lamb') & !str_detect(Ingredients, 'broth|salad') ~ 'lamb',
#     str_detect(Ingredients, 'pork|bacon|sausage|ham|salami') & !str_detect(Ingredients, 'lard|beef|burger') ~ 'pork',
#     str_detect(Ingredients, 'chicken|turkey|duck|goose') & !str_detect(Ingredients, 'broth|fat|sauce') ~ 'poultry',
# 
#     str_detect(Ingredients, 'pollock|cod|anglerfish|fish cakes coarse|fish ball|fish stick|haddock|grouper|catfish|sea bass|halibut|tuna') ~ 'lean fish', #The fishcakes are made out of mostly haddock
#     str_detect(Ingredients, 'salmon|trout|arctic char|mackerel|herring|sardine|anchovy') & !str_detect(Ingredients, 'roe') ~ 'oily fish',
#     str_detect(Ingredients, 'squid|prawn|shrimp|mussel|crab|lobster|shellfish|scampi|clam|scallop') & !str_detect(Ingredients, 'paste') ~ 'shellfish'
# 
#     )) %>%
#   #Vegetarian or vegan foods
#   group_by(recipe_name, foodgroup) %>%
#   mutate(
#     protein = case_when(
# 
#       !any(foodgroup %in% c('Seafood', 'Eggs', 'Milk and dairy products', 'Meat and nmeat products', 'Fish, seafood, amphibians, reptiles and invertebrates')) &
#         !str_detect(Ingredients, 'lamb|game|fish|beef|chicken|turkey|shrimp|cheese|condensed cream|egg\\b|duck|honey|pork|mayonnaise|barbeque|oyster|worcestershire|sausage|ham\\b|shortening|sour cream|tzatziki') &
#         !Ingredients %in% c('butter', 'unsalted butter', 'butter clarified ghee', 'butter for cooking',
#                             'buttermilk', 'refrigerated buttermilk biscuit dough', 'spice butter', 'puff pastry',
#                             'shop-bought shortcrust pastry') ~ 'plants',
#       !any(foodgroup %in% c('Seafood', 'Meat and\nmeat products')) & !str_detect(Ingredients, 'condensed cream|beef|fish|shrimp|duck|sausage|shortening') ~ 'vegetarian',
#       TRUE ~ protein),
#     vegetarian = case_when(
#       protein %in% c('plants', 'vegetarian') ~ 'vegetarian',
#       TRUE ~ protein
# 
#       )) %>% group_by(recipe_name) #%>%
#     #Clean up
#     # summarise(type = case_when(
#     #   any(protein == 'beef') ~ 'Beef',
#     #   any(protein == 'game') ~ 'Game',
#     #   any(protein == 'lamb') ~ 'Lamb',
#     #   any(protein == 'poultry') ~ 'Poultry',
#     #   any(protein == 'lean fish') ~ 'Lean fish',
#     #   any(protein == 'oily fish') ~ 'Oily fish',
#     #   any(protein == 'shellfish') ~ 'Shellfish',
#     #   any(protein == 'pork') ~ 'Pork',
#     #   all(protein == 'plants') ~ 'Vegan',
#     #   all(vegetarian == 'vegetarian') ~ 'Vegetarian'
#     #   )) %>% ungroup()

```

## Plot data
### Nutrient content
Looking at how much of the recommended intake of nutrient found per portion of the recipes.
```{r nutrients-rdi, fig.width=8, fig.height=5}
# Recommended intake of nutrients for adult women in Norway
nutrient_units <- tibble(
  'nutrient' = c('Vitamin A', 'Retinol', 'Beta-carotene',
                'Vitamin D','Vitamin E', 'Thiamin',
                'Riboflavin','Niacin', 'Vitamin B6',
                'Folate', 'Vitamin B12', 'Vitamin C',
                'Calcium', 'Iron', 'Sodium',
                'Potassium', 'Magnesium', 'Zinc',
                'Selenium', 'Copper', 'Phosphorus',
                'Iodine'),  
  'rdi' = c(700, 700, 8400,
            10, 8, 1.1,
            1.3, 15, 1.2,
            400, 2, 75,
            800, 15, 2300,
            3100, 280, 7,
            50, 0.9, 600,
            150))

#Plot function
buildPlot <- function(dataframe, title, x_axis, x_axis_title, y_axis, y_axis_title, ylim, title_text_size = 12) {
  ggplot(dataframe, aes(x = .data[[x_axis]], y = .data[[y_axis]])) +
    geom_half_boxplot(side = "r", outlier.shape = NA) +
    geom_half_violin() +
    geom_quasirandom(alpha = 0.5, size = 2, shape = 21) +
    labs(
      title = title,
      x = x_axis_title,
      y = y_axis_title
    ) +
    coord_cartesian(ylim = ylim) +
    theme(plot.title = element_text(size = title_text_size))
}

# Calculate pct of intake per portion
rdi_per_portion <- calculations$nutrient_content %>%
  right_join(nutrient_units) %>%
  mutate(pct_rdi_per_portion = (nutrient_content_per_portion/rdi)*100)

#Minerals
buildPlot(dataframe = rdi_per_portion %>% filter(nutrient %in% c(
  'Calcium', 'Iron', 'Sodium',
  'Potassium', 'Magnesium', 'Zinc',
  'Selenium', 'Copper', 'Phosphorus',
  'Iodine')), 
  title = "Percent of recommended daily intake of minerals found per portion of the recipes", x_axis = "nutrient", x_axis_title = "Mineral", y_axis = "pct_rdi_per_portion", y_axis_title = "Percent of RDI", ylim = c(0, 600)) #Cut out a few outliers to make plot more readable
# Water soluble vitamins
buildPlot(dataframe = rdi_per_portion %>% filter(nutrient %in% c(
  'Thiamin', 'Riboflavin','Niacin', 'Vitamin B6',
  'Folate', 'Vitamin B12', 'Vitamin C')), 
  title = "Percent of recommended daily intake of water soluble vitamins found per portion of the recipes", x_axis = "nutrient", x_axis_title = "Water soluble vitamin", y_axis = "pct_rdi_per_portion", y_axis_title = "Percent of RDI", ylim = c(0, 500))

# Fat soluble vitamins
buildPlot(dataframe = rdi_per_portion %>% filter(nutrient %in% c(
  'Vitamin A', 'Retinol', 'Beta-carotene', 'Vitamin D','Vitamin E')), 
  title = "Percent of recommended daily intake of fat soluble vitamins found per portion of the recipes", x_axis = "nutrient", x_axis_title = "Fat soluble vitamin", y_axis = "pct_rdi_per_portion", y_axis_title = "Percent of RDI", ylim = c(0, 750))
```

### Environmental impact of the recipes
Using ggplot to visualise the environmental impact of the recipes 
```{r nutrient-content-env-impact, fig.width=5, fig.height=5}
#Environmental impact per portion
buildPlot(dataframe = calculations$environmental_impact, 
  title = "Environmental impact per portion of Oda recipes SHARP ID", x_axis = "environmental_impact_indicator", x_axis_title = "Environmental impact indicator", y_axis = "environmental_impact_per_portion", y_axis_title = "Environmental impact in", ylim = c(0, 25), title_text_size = 11) +
  facet_wrap(~environmental_impact_indicator, scales = "free") 
# Recipes with only one portion have a larger environmental impact than recipes with more than 1 portion, look only at those.
buildPlot(dataframe = calculations$environmental_impact %>% filter(number_of_portions != 1), 
  title = "Environmental impact per portion of Oda recipes \nwith number of portions >1 mapped to SHARP ID", x_axis = "environmental_impact_indicator", x_axis_title = "Environmental impact indicator", y_axis = "environmental_impact_per_portion", y_axis_title = "Environmental impact", ylim = c(0, 7), title_text_size = 11) +
  facet_wrap(~environmental_impact_indicator, scales = "free") 
```
### Nutriscore
The Nutriscore is a front of pack label that should give consumers at-a-glance information about a foods healthiness. The Nutriscore is calculated per 100 g, from the amount of energy, macronutrients and percentage of fruit/vegetables and healthy oils in a food.
```{r nutriscore-function}
#Function to calculate the Nutriscore of a recipe
#' @param df Dataframe with nutrient content per 100g and percentage of weight of fruit/vegetables and healthy oils
calculateNutriscore <- function(df, raw_scores = FALSE){
  
  #Join them together and calculate nutriscore
  nutriscore_raw <- df %>%
    
    #Calculate nutriscore
    mutate(nutriscore_raw = case_when(
      #Energy
      nutrient == 'Kilojoules' & nutrient_content_per_hektogram <= 335 ~ 0,
      nutrient == 'Kilojoules' & (nutrient_content_per_hektogram >335 & nutrient_content_per_hektogram <= 670) ~ 1,
      nutrient == 'Kilojoules' & (nutrient_content_per_hektogram >670 & nutrient_content_per_hektogram <= 1005) ~ 2,
      nutrient == 'Kilojoules' & (nutrient_content_per_hektogram >1005 & nutrient_content_per_hektogram <= 1340) ~ 3,
      nutrient == 'Kilojoules' & (nutrient_content_per_hektogram >1340 & nutrient_content_per_hektogram <= 1675) ~ 4,
      nutrient == 'Kilojoules' & (nutrient_content_per_hektogram >1675 & nutrient_content_per_hektogram <= 2010) ~ 5,
      nutrient == 'Kilojoules' & (nutrient_content_per_hektogram >2010 & nutrient_content_per_hektogram <= 2345) ~ 6,
      nutrient == 'Kilojoules' & (nutrient_content_per_hektogram >2345 & nutrient_content_per_hektogram <= 2680) ~ 7,
      nutrient == 'Kilojoules' & (nutrient_content_per_hektogram >2680 & nutrient_content_per_hektogram <= 3015) ~ 8,
      nutrient == 'Kilojoules' & (nutrient_content_per_hektogram >3015 & nutrient_content_per_hektogram <= 3350) ~ 9,
      nutrient == 'Kilojoules' & nutrient_content_per_hektogram >3350 ~ 10,
      
      #(Added) Sugars
      nutrient == 'Sugar' & nutrient_content_per_hektogram <= 4.5 ~ 0,
      nutrient == 'Sugar' & (nutrient_content_per_hektogram >4.5 & nutrient_content_per_hektogram <= 9) ~ 1,
      nutrient == 'Sugar' & (nutrient_content_per_hektogram >9 & nutrient_content_per_hektogram <= 13.5) ~ 2,
      nutrient == 'Sugar' & (nutrient_content_per_hektogram >13.5 & nutrient_content_per_hektogram <= 18) ~ 3,
      nutrient == 'Sugar' & (nutrient_content_per_hektogram >18 & nutrient_content_per_hektogram <= 22.5) ~ 4,
      nutrient == 'Sugar' & (nutrient_content_per_hektogram >22.5 & nutrient_content_per_hektogram <= 27) ~ 5,
      nutrient == 'Sugar' & (nutrient_content_per_hektogram >27 & nutrient_content_per_hektogram <= 31) ~ 6,
      nutrient == 'Sugar' & (nutrient_content_per_hektogram >31 & nutrient_content_per_hektogram <= 36) ~ 7,
      nutrient == 'Sugar' & (nutrient_content_per_hektogram >36 & nutrient_content_per_hektogram <= 40) ~ 8,
      nutrient == 'Sugar' & (nutrient_content_per_hektogram >40 & nutrient_content_per_hektogram <= 45) ~ 9,
      nutrient == 'Sugar' & nutrient_content_per_hektogram >45 ~ 10,
      
      #Saturated Fat
      nutrient == 'SatFa' & nutrient_content_per_hektogram <= 1 ~ 0,
      nutrient == 'SatFa' & (nutrient_content_per_hektogram >1 & nutrient_content_per_hektogram <= 2) ~ 1,
      nutrient == 'SatFa' & (nutrient_content_per_hektogram >2 & nutrient_content_per_hektogram <= 3) ~ 2,
      nutrient == 'SatFa' & (nutrient_content_per_hektogram >3 & nutrient_content_per_hektogram <= 4) ~ 3,
      nutrient == 'SatFa' & (nutrient_content_per_hektogram >4 & nutrient_content_per_hektogram <= 5) ~ 4,
      nutrient == 'SatFa' & (nutrient_content_per_hektogram >5 & nutrient_content_per_hektogram <= 6) ~ 5,
      nutrient == 'SatFa' & (nutrient_content_per_hektogram >6 & nutrient_content_per_hektogram <= 7) ~ 6,
      nutrient == 'SatFa' & (nutrient_content_per_hektogram >7 & nutrient_content_per_hektogram <= 8) ~ 7,
      nutrient == 'SatFa' & (nutrient_content_per_hektogram >8 & nutrient_content_per_hektogram <= 9) ~ 8,
      nutrient == 'SatFa' & (nutrient_content_per_hektogram >9 & nutrient_content_per_hektogram <= 10) ~ 9,
      nutrient == 'SatFa' & nutrient_content_per_hektogram >10 ~ 10,
      
      #Sodium
      nutrient == 'Sodium' & nutrient_content_per_hektogram <= 90 ~ 0,
      nutrient == 'Sodium' & (nutrient_content_per_hektogram >90 & nutrient_content_per_hektogram <= 180) ~ 1,
      nutrient == 'Sodium' & (nutrient_content_per_hektogram >180 & nutrient_content_per_hektogram <= 270) ~ 2,
      nutrient == 'Sodium' & (nutrient_content_per_hektogram >270 & nutrient_content_per_hektogram <= 360) ~ 3,
      nutrient == 'Sodium' & (nutrient_content_per_hektogram >360 & nutrient_content_per_hektogram <= 450) ~ 4,
      nutrient == 'Sodium' & (nutrient_content_per_hektogram >450 & nutrient_content_per_hektogram <= 540) ~ 5,
      nutrient == 'Sodium' & (nutrient_content_per_hektogram >540 & nutrient_content_per_hektogram <= 630) ~ 6,
      nutrient == 'Sodium' & (nutrient_content_per_hektogram >630 & nutrient_content_per_hektogram <= 720) ~ 7,
      nutrient == 'Sodium' & (nutrient_content_per_hektogram >720 & nutrient_content_per_hektogram <= 810) ~ 8,
      nutrient == 'Sodium' & (nutrient_content_per_hektogram >810 & nutrient_content_per_hektogram <= 900) ~ 9,
      nutrient == 'Sodium' & nutrient_content_per_hektogram >900 ~ 10,
      
      #Fruits and veggies an legumes and nuts and (certain) oils
      nutrient == 'nutriscore_amounts_pct' & nutrient_content_per_hektogram <= 40 ~ 0,
      nutrient == 'nutriscore_amounts_pct' & (nutrient_content_per_hektogram > 40 & nutrient_content_per_hektogram <= 60) ~ 1,
      nutrient == 'nutriscore_amounts_pct' & (nutrient_content_per_hektogram > 60 & nutrient_content_per_hektogram <= 80) ~ 2,
      nutrient == 'nutriscore_amounts_pct' & nutrient_content_per_hektogram > 80 ~ 5,
      
      #Fibre
      nutrient == 'Dietary fibre' & nutrient_content_per_hektogram <= 0.9 ~ 0,
      nutrient == 'Dietary fibre' & (nutrient_content_per_hektogram >0.9 & nutrient_content_per_hektogram <=1.9) ~ 1,
      nutrient == 'Dietary fibre' & (nutrient_content_per_hektogram >1.9 & nutrient_content_per_hektogram <=2.8) ~ 2,
      nutrient == 'Dietary fibre' & (nutrient_content_per_hektogram >2.8 & nutrient_content_per_hektogram <=3.7) ~ 3,
      nutrient == 'Dietary fibre' & (nutrient_content_per_hektogram >3.7 & nutrient_content_per_hektogram <=4.7) ~ 4,
      nutrient == 'Dietary fibre' & nutrient_content_per_hektogram >4.7 ~ 5,
      
      #Protein
      nutrient == 'Protein' & nutrient_content_per_hektogram <= 1.6 ~ 0,
      nutrient == 'Protein' & (nutrient_content_per_hektogram >1.6 & nutrient_content_per_hektogram <=3.2) ~ 1,
      nutrient == 'Protein' & (nutrient_content_per_hektogram >3.2 & nutrient_content_per_hektogram <=4.8) ~ 2,
      nutrient == 'Protein' & (nutrient_content_per_hektogram >4.8 & nutrient_content_per_hektogram <=6.4) ~ 3,
      nutrient == 'Protein' & (nutrient_content_per_hektogram >6.4 & nutrient_content_per_hektogram <=8.0) ~ 4,
      nutrient == 'Protein' & nutrient_content_per_hektogram >8.0 ~ 5,
    )) %>% drop_na(nutriscore_raw) #Drop nutrients not needed for calculations
  
  #temporary categories used to calculate final score
  categories <- nutriscore_raw %>%
    mutate(category = case_when(
      nutrient %in% c('Kilojoules', 'Sugar', 'SatFa', 'Sodium') ~ 'N',
      nutrient %in% c('nutriscore_amounts_pct', 'Dietary fibre', 'Protein') ~ 'P'
    )) %>% drop_na(category) %>%
    #get scores for the categories
    group_by(recipe_name, category) %>%
    summarise(nutriscore_raw = sum(nutriscore_raw)) %>%
    ungroup() %>%
    #Turn into wide format
    pivot_wider(
      names_from = 'category',
      values_from = 'nutriscore_raw'
    )
  
  categories_2 <- nutriscore_raw %>%
    mutate(category = case_when(
      nutrient == 'nutriscore_amounts_pct' ~ 'FruitVegLegumesNutsOils',
      nutrient == 'Dietary fibre' ~ 'Fibre',
      nutrient == 'Protein' ~ 'Protein'
    )) %>% drop_na(category) %>%
    #get scores for the categories
    group_by(recipe_name, category) %>%
    summarise(nutriscore_raw = sum(nutriscore_raw)) %>%
    ungroup() %>%
    #Turn into wide format
    pivot_wider(
      names_from = 'category',
      values_from = 'nutriscore_raw'
    ) %>%
    # 0 for recipes without fruit legumes nuts
    mutate(FruitVegLegumesNutsOils = case_when(
      is.na(FruitVegLegumesNutsOils) ~ 0,
      TRUE ~ FruitVegLegumesNutsOils
    ))
  
  #Calculate nutriscore 
  nutriscore <- full_join(categories, categories_2) %>%
    mutate(nutriscore = case_when(
      N < 11 ~ N-P,
      N >= 11 & FruitVegLegumesNutsOils >= 5 ~ N-P,
      N >= 11 & FruitVegLegumesNutsOils < 5 ~ N-(FruitVegLegumesNutsOils+Fibre)
    ),
    nutriscore_letter = case_when(
      nutriscore <= -1 ~ 'A',
      nutriscore >= 0 & nutriscore <=2 ~ 'B',
      nutriscore >= 3 & nutriscore <=10 ~ 'C',
      nutriscore >= 11 & nutriscore <=18 ~ 'D',
      nutriscore >= 19 ~ 'E',
    )) %>%
    #Turn letters into factor
    mutate(nutriscore_letter = as.factor(nutriscore_letter))
}

```
```{r nutriscore-dataframe, include = FALSE}
calculations$nutrient_content_100g <- mapped_to_database$nutrients %>%
  calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g")
```
```{r nutriscore-calculation}
# Calculate percentage of fruit/vegetables/oil in a recipe
# Weight of all ingredients, use ingredients mapped to SHARP db to use the foodgroups in SHARP later
nutriscore <- mapped_to_database$environmental_impact %>%
  #Only look at recipes that have no composite ingredients for simplicity's sake
  group_by(recipe_name) %>% filter(!any(database_reference == "Not in database")) %>%
  #Use foodgroups from SHARP
  left_join(., SHARP2018_foodgroups) %>%
  select(recipe_name, Ingredients, Amounts, foodgroup) %>% distinct() %>%
  # #Mark the foods relevant for Nutriscore
  mutate(
     nutriscore_foods = case_when(
       (foodgroup %in% c('Vegetables and vegetable products', 'Fruit and fruit products', 'Fruit/vegetable juice and nectar', 'Legumes, nuts, oilseeds and spices') & !str_detect(Ingredients, 'mushroom|coconut milk|sesame seed|pine')) | str_detect(Ingredients, 'olive oil|rapeseed oil|walnut oil') ~ 'nutriscore_fruit_veg_legumes_nuts_oils',
       TRUE ~ "Other")
    ) %>%
  # Calculate percent of total recipe weight
  #Dried fruit/veg and ketchup/purees should be doubled
  mutate(
      Amounts = case_when(
        nutriscore_foods == "nutriscore_fruit_veg_legumes_nuts_oils" & str_detect(Ingredients, 'paste tomato|ketchup|dried|raisin|prune') ~ Amounts*2,
        TRUE ~ Amounts)) %>%
  group_by(recipe_name, nutriscore_foods) %>%
  mutate(weight_foodgroups = sum(Amounts, na.rm = TRUE)) %>% ungroup() %>%
  select(recipe_name, weight_foodgroups, nutriscore_foods) %>% distinct() %>%
  group_by(recipe_name) %>%
  #Use same column name as nutrient content dataframe
  mutate(
    nutrient_content_per_hektogram = (weight_foodgroups/sum(weight_foodgroups, na.rm = TRUE))*100,
         nutrient = "nutriscore_amounts_pct") %>%
  filter(nutriscore_foods != "Other") %>%
  select(recipe_name, nutrient, nutrient_content_per_hektogram) %>%
  #Add nutrients and calculate Nutriscore
  bind_rows(., calculations$nutrient_content_100g) %>%
  calculateNutriscore()
```
Plotting the results
```{r nutriscore-plots}
ggplot(nutriscore, aes(x = nutriscore_letter, fill = nutriscore_letter)) +
  geom_bar(show.legend = FALSE) +
  labs(
    title = "Nutriscore of a subset of Oda recipes",
    y = "Number of recipes",
    x = "Nutriscore letter"
  ) +
  #Use Nutriscore coloring scheme
  scale_fill_manual(values = c("A" = "#028141", "B" = "#85BB2F", "C" = "#FECB03", "D" = "#EE8100", "E" = "#E63D11"))

```
