---
title: "Map foods to database"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Map foods to database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Load libraries
```{r setup, message = FALSE}
library(sustainableNutRients)
library(tidyverse)
```

# Read in some data
sustainableRecipes has a dataset with 100 recipes scraped from the Norwegian online store Oda that can be used to check out the different functions. For more info, read the documentation for the dataset with ?recipes_oda.
```{r read-data}
# Read the data
data("recipes_oda")

# What it looks like. The "ingredient_is_part_of" column show which part of the recipe the ingredient comes from
# such as marinade, sauce, side or main.
head(recipes_oda$recipes %>% select(-ingredient_is_part_of))
```

# Standardise food names and units before mapping to nutrient and sustainability databases
Before food items can be mapped to the databases, food names and units must be standardised. This is done by using the function standardiseFoodList.
```{r standardise-recipes}
# Run standardiseFoodList() function.
standardised_recipes <- recipes_oda$recipes %>%
  standardiseFoodList()
```
While the funtion recognizes many different food items, it is advisable to do a quick check of the standardisations and fix potential errors. The function checkFoodStandardisation provides a list of the original ingredient names next to their standardised names.
```{r check-standardisation}
# Check food standardisation
standardisation_check <- checkFoodStandardisation(standardised_recipes)

head(standardisation_check)
```

# Map to database(s)
After foods and amounts have been standardised, foods can be mapped to the databases for volume/weight conversion, nutrient content and environmental impact calculations.

## Volume weight convertion
Nutrient content and environmental impact is calculated per 100 g of a food, so first all amounts have to be turned into hektograms. This is done by first mapping the list of foods and their amounts to the volume_weight database.

As with standardisation, the database mapping can be checked using the function checkFoodDatabaseMapping. This information can then be used to create new values that can be added to the database mapping function using the "additional_entries" argument. For the recipes from Oda, there are a few ingredients that are not found in the volume_weight database so they are added by creating additional entries with the createNewDataBaseEntry function.

```{r volume-weight-convertion, message = FALSE, results = FALSE}
#Some ingredients in the recipes are not found in the database and have been mapped to a similar but wrong ingredient. These will be added by creating additional entries to the database
not_in_volume_weight <- tibble(
  tmp = c("bolognese_vegetarian pcs 370",
          "hummus pcs 165",
          "pizza_dough pcs 210",
          "sauce_pizza pack 400",
          "sauce_pad pcs 150",
          "soup_tomato pcs 530",
          "vanilla_pod pcs 3")
) %>% separate(col = tmp, into = c("recipe_name", "unit", "grams_per_unit"), sep = " ") %>%
  mutate(grams_per_unit = as.numeric(grams_per_unit)) %>%
  createNewDatabaseEntry(., "volume_weight")

#Map to volume_weight database
volume_weight <- findFoodInDatabase(standardised_recipes, "volume_weight", additional_entries = not_in_volume_weight)

#Calculate weight of ingredients
weighted_ingredients <- calculateWeightOfIngredients(volume_weight)
```

## Nutrient and/or environmental indicators
With the weight of our foods, their nutrient content and environmental impact can be calculated! This is similar to converting volume to weight units, by using the findFoodInDatabase function. To add additional entries, for example of composite foods, a recipe for these foods must first be calculated (as shown below, with a shortened number of composite ingredients, a larger list can be found in data(composite_ingredients_oda)) and then additional_entries to the findFoodInDatabase function can be created by createNewDatabaseEntry using the "nutrient" or "sustainability" argument depending on which database is used.

```{r nutrient-composite-ingredients-show, message = FALSE, results=FALSE}
nutrients_additional_ingredients <- tibble(
  #Find recipes of the composite ingredients
  tmp = c(
    #From Oda webstore
    "raw vegetables_assorted;34 g carrot
    33 g parsnip
    33 g swede",
    #From https://cooking.nytimes.com/recipes/1022154-vegetarian-bolognese
    "bolognese_vegetarian;3 tablespoons unsalted butter
    2 tablespoons extra-virgin olive oil
    1 large yellow onion, roughly chopped into ¼-inch pieces
    2 medium carrots, roughly chopped into ¼-inch pieces
    1.25 pound cauliflower
    0.75 cup tomato paste
    2 tablespoons low-sodium soy sauce
    4 garlic cloves, coarsely chopped
    0.75 cup whole milk
    1 fresh bay leaf or thyme sprig
    1 pound rigatoni or another ridged dried pasta, or fresh pappardelle or tagliatelle
    0.5 cup finely grated Parmesan, plus more for serving"
  )) %>%
  #Format
  separate(., tmp, into = c("recipe_name", "Ingredients"), sep = ";") %>%
  separate_rows(., Ingredients, sep = "\\n") %>%
  #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
  standardiseFoodList() %>%
  findFoodInDatabase(., "volume_weight") %>%
  calculateWeightOfIngredients() %>%
  findFoodInDatabase(., "nutrients") %>%
  #Calculate nutrients per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
  calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g") %>%
  #Format as new database entries for the nutrients database
  createNewDatabaseEntry(., "nutrients")
```
```{r nutrient-composite-ingredients-hide, message = FALSE, echo=FALSE, results=FALSE, warning=FALSE}
#Read in full list of composite ingredients in these recipes
data(composite_ingredients_oda)
#Calculate nutrient content
nutrients_additional_ingredients <- composite_ingredients_oda %>%
  #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
  standardiseFoodList() %>%
  findFoodInDatabase(., "volume_weight") %>%
  calculateWeightOfIngredients() %>%
  findFoodInDatabase(., "nutrients") %>%
  #Calculate nutrients per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
  calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g") %>%
  #Format as new database entries for the nutrients database
  createNewDatabaseEntry(., "nutrients")
```
The output from calculateWeightOfIngredients is used to calculate total nutrient content (default of calculateNutrientConcentOfFoodlist) of the recipes:
```{r nutrient-environmental-calculations_total, message=FALSE, results = FALSE}
nutrient_content <- findFoodInDatabase(
  weighted_ingredients, "nutrients", 
  #Using the additional entries to the database calculated above
  additional_entries = nutrients_additional_ingredients) %>%
  calculateNutrientContentOfFoodlist()
```
Instead of calculating nutrient content for the total nutrient content of a recipe, it can be calculated "per portion" or "per 100 g" by using the calculate_nutrient argument. In the same manner, environmental impact can be calculated using calculateEnvironmentalImpactOfFoodlist that has the same arguments that calculateNutrientContentOfFoodlist.

