---
title: "Map foods to database"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Map foods to database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Load libraries
```{r setup, message = FALSE}
library(sustainableNutRients)
library(tidyverse)
```

# Read in some data
sustainableRecipes has a dataset with 100 recipes scraped from the Norwegian online store Oda that can be used to check out the different functions. For more info, read the documentation for the dataset with ?recipes_oda.
```{r read-data}
# Read the data
data("recipes_oda")

# What it looks like. The "ingredient_is_part_of" column show which part of the recipe the ingredient comes from
# such as marinade, sauce, side or main.
head(recipes_oda$recipes %>% select(-ingredient_is_part_of))
```

# Standardise food names and units before mapping to nutrient and sustainability databases
Before food items can be mapped to the databases, food names and units must be standardised. This is done by using the function standardiseFoodList.
```{r standardise-recipes}
# Run standardiseFoodList() function.
standardised_recipes <- recipes_oda$recipes %>%
  standardiseFoodList()
```
While the funtion recognizes many different food items, it is advisable to do a quick check of the standardisations and fix potential errors. The function checkFoodStandardisation provides a list of the original ingredient names next to their standardised names.
```{r check-standardisation}
# Check food standardisation
standardisation_check <- checkFoodStandardisation(standardised_recipes)

head(standardisation_check)
```

# Map to database(s)
After foods and amounts have been standardised, foods can be mapped to the databases for volume/weight conversion, nutrient content and environmental impact calculations.

## Volume weight convertion
Nutrient content and environmental impact is calculated per 100 g of a food, so first all amounts have to be turned into hektograms. This is done by first mapping the list of foods and their amounts to the volume_weight database.

As with standardisation, the database mapping can be checked using the function checkFoodDatabaseMapping. This information can then be used to create new values that can be added to the database mapping function using the "additional_entries" argument. For the recipes from Oda, there are a few ingredients that are not found in the volume_weight database so they are added by creating additional entries with the createNewDataBaseEntry function.

```{r volume-weight-convertion, message = FALSE, results = FALSE}
#Some ingredients in the recipes are not found in the database and have been mapped to a similar but wrong ingredient. These will be added by creating additional entries to the database
not_in_volume_weight <- tibble(
  tmp = c("bolognese_vegetarian pcs 370",
          "hummus pcs 165",
          "pizza_dough pcs 210",
          "sauce_pizza pack 400",
          "sauce_pad pcs 150",
          "soup_tomato pcs 530",
          "vanilla_pod pcs 3")
) %>% separate(col = tmp, into = c("Ingredients", "unit", "grams_per_unit"), sep = " ") %>%
  mutate(grams_per_unit = as.numeric(grams_per_unit)) %>%
  createNewDatabaseEntry(., "volume_weight")

#Map to volume_weight database
volume_weight <- findFoodInDatabase(standardised_recipes, "volume_weight", additional_entries = not_in_volume_weight)

#Calculate weight of ingredients
weighted_ingredients <- calculateWeightOfIngredients(volume_weight)
```

## Nutrient and/or environmental indicators
With the weight of our foods, their nutrient content and environmental impact can be calculated! This is similar to converting volume to weight units, by using the findFoodInDatabase function. To add additional entries, for example of composite foods, a recipe for these foods must first be calculated (as shown below, with a shortened number of composite ingredients) and then additional_entries to the findFoodInDatabase function can be created by createNewDatabaseEntry using the "nutrient" or "sustainability" argument.

```{r nutrient-composite-ingredients-show, message = FALSE, results=FALSE}
nutrients_additional_ingredients <- tibble(
  
  #Find recipes of the composite ingredients
  tmp = c(
    #From Oda webstore
    "raw vegetables_assorted;34 g carrot
    33 g parsnip
    33 g swede",
    #From https://cooking.nytimes.com/recipes/1022154-vegetarian-bolognese
    "bolognese_vegetarian;3 tablespoons unsalted butter
    2 tablespoons extra-virgin olive oil
    1 large yellow onion, roughly chopped into ¼-inch pieces
    2 medium carrots, roughly chopped into ¼-inch pieces
    1.25 pound cauliflower
    0.75 cup tomato paste
    2 tablespoons low-sodium soy sauce
    4 garlic cloves, coarsely chopped
    0.75 cup whole milk
    1 fresh bay leaf or thyme sprig
    1 pound rigatoni or another ridged dried pasta, or fresh pappardelle or tagliatelle
    0.5 cup finely grated Parmesan, plus more for serving"
  )) %>%
  #Format
  separate(., tmp, into = c("recipe_name", "Ingredients"), sep = ";") %>%
  separate_rows(., Ingredients, sep = "\\n") %>%
  #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
  standardiseFoodList() %>%
  findFoodInDatabase(., "volume_weight") %>%
  calculateWeightOfIngredients() %>%
  findFoodInDatabase(., "nutrients") %>%
  #Calculate nutrients per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
  calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g") %>%
  #Format as new database entries for the nutrients database
  createNewDatabaseEntry(., "nutrients")
```
```{r nutrient-composite-ingredients-hide, message = FALSE, echo=FALSE, results=FALSE, warning=FALSE}
nutrients_additional_ingredients <- tibble(
  
  #Find recipes of the composite ingredients
  tmp = c(
    #From Oda webstore
    "raw vegetables_assorted;34 g carrot
    33 g parsnip
    33 g swede",
    #From https://cooking.nytimes.com/recipes/1022154-vegetarian-bolognese
    "bolognese_vegetarian;3 tablespoons unsalted butter
    2 tablespoons extra-virgin olive oil
    1 large yellow onion, roughly chopped into ¼-inch pieces
    2 medium carrots, roughly chopped into ¼-inch pieces
    1.25 pound cauliflower
    0.75 cup tomato paste
    2 tablespoons low-sodium soy sauce
    4 garlic cloves, coarsely chopped
    0.75 cup whole milk
    1 fresh bay leaf or thyme sprig
    1 pound rigatoni or another ridged dried pasta, or fresh pappardelle or tagliatelle
    0.5 cup finely grated Parmesan, plus more for serving",
    #Lapskausblanding, oda
    "vegetable mix_frozen;63 g potato
    20 g carrot
    9 g celeriac
    8 g leek",
    #Inspiredtaste.net hummus
    "hummus;250 g chickpeas
    60 ml lemon juice
    60 ml tahini
    1 clove garlic
    30 ml olive oil
    0.5 tsp cumin
    2.5 tbsp water",
    #Inspredtaste.net coleslaw
    "coleslaw;2 pound cabbage
    2 carrots
    0.5 cup fresh parsley
    170 g mayonnaise
    2 tbsp apple cider vinegar
    2 tbsp mustard
    0.25 tsp salt
    1.5 tsp sugar",
    #Cookie and Kate tzatziki
    "tzatziki;10 ounce cucumber
    1.5 cup plain greek yoghurt
    2 tbsp olive oil
    2 tbsp fresh dill
    1 tbsp lemon juice
    1 clove garlic
    0.5 tsp sea salt",
    #Minimalist baker chana masala
    "chana_masala;3 tbsp coconut oil
    1 onion
    1 tbsp ground cumin
    0.75 tsp salt
    6 clove garlic
    2 tbsp fresh ginger
    0.5 cup cilantro
    2.5 pcs green chili
    1 tbsp ground coriander
    1 tsp chili powder
    1 tsp ground turmeric
    28 ounce tomatoes
    15 ounce canned chickpeas
    1 tsp garam masala
    2.5 tsp coconut sugar
    2 tbsp lemon juice",
    #Simply Recipe's lasagna
    "lasagna;2 tsp olive oil
    1 pound beef chuck
    0.5 pcs onion
    0.5 pcs bell pepper
    2 clove garlic
    28 ounce tomato sauce
    3 ounce tomato paste
    14 ounce tomatoes
    2 tsp dried oregano
    0.25 cup parsley
    1 tbsp italian seasoning
    1 pinch garlic powder
    1 tbsp red wine vinegar
    2 tbsp sugar
    0.5 pound pasta
    15 ounce ricotta
    11.5 pound mozzarella cheese
    0.25 pound parmesan cheese",
    #Oat cream recipe
    "cream_plant-based;2 dl oat milk
    1 dl rapeseed oil",
    #The edgy veg Vegan minced meat
    "plant based_minced meat;1 cup walnuts
    16 oz mushrooms
    3 tbsp olive oil
    2 tbsp soy sauce
    1 tsp ground cumin
    2 tsp ground coriander
    0.5 tsp salt
    1 tsp white pepper
    2 tsp smoked paprika powder
    2 tsp garlic powder
    1.5 tsp onion powder
    0.25 tsp chili powder",
    #Oda smoothie mix
    "smoothie mix_tropical;35 g pineapple
    35 g mango
    30 g banana",
    #The mediterraneandish falafels
    "falafel;2 cup dried chickpeas
    0.5 tsp baking soda
    1 cup fresh parsley
    0.75 cup fresh cilantro
    0.5 cup fresh dill
    1 onion
    7.5 clove garlic
    1 tbsp ground black pepper
    1 tbsp ground cumin
    1 tsp cayenne pepper
    1 tsp baking powder
    2 tbsp sesame seeds",
    #Thai soup for shrimp
    "soup_thai;2 tbsp butter
    2 clove garlic
    1 onion
    1 red bell pepper
    1 tbsp fresh ginger
    2 tbsp curry paste
    12 ounce coconut milk
    3 cup chicken stock
    1 lime the juice of
    2 tbsp fresh cilantro",
    #tine porridge
    "porridge;2 dl rice
    4 dl water
    10 dl whole milk
    1 ts salt
    2 tbsp butter
    1 dl sugar
    2 tbsp cinnamon",
    #Baker betty cookies
    "cookie_cookie dough;113 g butter
    150 g brown sugar
    1 egg
    1 tsp vanilla extract
    0.5 tsp baking powder
    0.5 tsp salt
    150 g all purpose flour
    150 g semi-sweet chocolate chips",
    #The garden grazer mexican lentil bean salad
    "bean salad_mexican lentil;1 cup green lentils dry
    15 ounce can black beans
    1 red bell pepper
    0.33 cup onion
    6 oz cherry tomatoes
    0.67 cup cilantro
    2 tbsp lime juice
    1 tsp dijon mustard
    1.5 clove garlic
    0.5 tsp ground cumin
    0.5 tsp dried oregano
    0.25 tsp salt",
    #The recipe well bean and olive salad
    "bean salad;19 oz cannellini beans canned
    19 oz kidney beans canned
    15 oz can kidney beans
    0.5 pcs cucumber
    2 cup cherry tomatoes
    1 bell pepper
    0.5 cup black olives
    0.5 cup fresh parsley
    0.5 cup red onion
    2 tbsp fresh basil
    0.75 cup feta cheese",
    #Joyfood sunshine chocolate chip cookies
    "biscuit_chocolate chip;1 cup butter
    1 cup white sugar
    1 cup brown sugar
    2 tsp vanilla extract
    2 eggs
    3 cup all purpose flour
    1 tsp baking soda
    1 tsp salt
    12 oz chocolate chips",
    #Sugar salt magic homemade oreos
    "biscuit_oreo;130 g flour
    35 g corn starch
    0.33 cup cocoa
    0.5 tsp baking soda
    95 g unsalted butter
    1 tbsp milk
    113 g unsalted butter
    1.5 cup confectioners sugar
    1 tsp vanilla extract",
    #Tasted recipes homemade salted biscuits
    "biscuit_salt; 60 g butter
    50 g sugar
    1 tbs sesame seed
    1 tsp baking soda
    136 g all purpose flour
    1 tsp salt
    4.5 tbsp milk",
    #Damn delicious sourdough biscuits, sourdough starter from the clever carrpt
    "biscuit_sourdough;2 cup all purpose flour
    2 tsp sugar
    2 tsp baking powder
    1 tsp salt
    0.75 tsp baking soda
    0.5 cup unsalted butter
    0.5 cup buttermilk
    60 g whole wheat flour
    60 g water",
    #Becky's mindful kitchen Spelt biscuits
    "biscuit_spelt;2 cup spelt flour
    1.5 tsp salt
    1 tsp baking powder
    0.25 cup olive oil
    0.85 cup water",
    #Dinner then desert sweet and sour chicken
    "chicken_sweet and sour;1.5 pound chicken breast
    0.5 cup cornstarch
    2 eggs
    0.25 cup flour
    1 cup pineapple
    1 red bell pepper
    1 green bell pepper
    0.5 pcs onion
    0.5 cup sugar
    0.25 cup brown sugar
    0.33 cup ketchup
    4 tsp soy sauce
    2 clove garlic",
    #Cookie and cate chili sin carne
    "chili_sin carne;2 tbsp olive oil
    1 onion
    1 red bell pepper
    2 carrots
    2 stalk celery
    0.5 tsp salt
    4 clove garlic
    2 tbsp chili powder
    2 tsp cumin
    1.5 tsp smoked paprika powder
    1 tsp dried oregano
    28 ounce canned tomatoes
    30 ounce canned black beans
    15 ounce canned pinto beans
    2 cups water
    1 bay leaf
    2 tbsp fresh cilantro
    1.5 tsp lime juice",
    #Homemade home chocolate mousse miz
    "chocolate_mousse powder;1.25 cup sugar
    1 cup cornstarch
    1 cup milk powder
    0.25 cup cocoa powder
    0.5 pcs vanilla bean",
    #Chocolate covered cate chocolate spread
    "chocolate_spread; 240 g hazelnuts
    1.5 tbsp vanilla extract
    0.25 cup cocoa powder
    0.33 cup honey
    0.25 tsp salt
    0.5 cup milk
    2 tbsp coconut oil",
    #31 daily oat crackers
    "cracker_oat;0.5 cup walnuts
    1.75 cup rolled oats
    1 tsp salt
    1 tsp baking powder
    2.25 tsp italian seasoning
    0.25 tsp garlic powder
    1 cup milk",
    #Meny fish balls
    "fish_ball;600 g cod
    0.5 tsp salt
    1 tbsp potato starch
    1 egg
    4 dl whole milk",
    #BBC good food Fish cakes
    "fish_cake;450 g cod;
    2 leaf bay leaf
    150 ml milk
    350 g potatoes
    0.5 tsp lemon zest
    1 tbsp chpped parsley
    1 tbsp chives
    1 egg
    85 g breadcrumbs
    3.5 tbsp sunflower oil",
    #Crnuchy creamy sweet fish sticks
    "fish_stick;2 fillet cod
    0.5 tsp salt
    0.25 tsp black pepper
    1 egg
    2 tbsp cream
    0.25 cup all purpose flour
    0.33 cup breadcrumbs
    0.25 tså garlic powder
    0.25 tsp italian seasoning
    0.5 cup vegetable oil",
    #Ricardo cuisine meatballs in tomato sauce
    "meatball_tomato sauce;3 clove garlic
    1 leaf bay
    0.25 tsp red pepper flakes
    30 ml olive oil
    28 oz tomatoes
    2 slice white bread
    60 ml milk
    450 g ground beef
    150 g parmesan cheese
    1 egg
    1 clove garlic
    10 g chopped parsley
    0.5 tsp salt
    0.5 tsp ground fennel
    0.25 tsp dried oregano
    0.25 tsp red pepper flakes",
    #BBC good food ravioli
    "pasta_filled;300 g all purpose flour
    2 eggs
    2 egg yolks
    1 tbsp olive oil
    2 clove garlic
    200 g baby spinach
    0.5 lemon zested
    250 g ricotta cheese",
    #Oda curry paste
    "curry_paste;3 tbsp coriander seed
    1 tsp fennel seed
    1 tsp pepper
    1 tsp tumeric
    2 tbsp cumin
    1 tsp cinnamon
    1 tsp paprika powder
    1 tsp salt
    1 pcs chili pepper
    4 clove garlic
    1 tsp fresh ginger
    4 tbsb white wine vinegar
    1 tbsp tomato puree",
    #Matprat langpanne Pizza base
    "pizza_base;3 dl water
    0.5 pack yeast
    2 tbsp oil
    0.5 tsp salt
    7 dl all purpose flour",
    #Also use as pizza dough
    "pizza_dough;3 dl water
    0.5 pack yeast
    2 tbsp oil
    0.5 tsp salt
    7 dl all purpose flour",
    #Matprat crem potatoes
    "potato_cream;4 dl cream
    1 clove garlic
    0.5 tsp salt
    0.25 tsp pepper
    1 kg potato
    1 tbsp butter
    1 dl grated cheese",
    #Matprat potato mash
    "potato_mash;10 potatoes
    2 dl milk
    2 tbsp butter
    0.5 tsp salt
    0.5 tsp pepper
    0.25 tsp nutmeg",
    #Klikk salt sticks
    "salt_stick; 4 dl wheat flour
    1.5 dl water
    10 g yeast dry
    0.5 tsp salt",
    #BBC goodfood vegetable spread
    "vegetable_spread;390 g chickpeas
    1 clove garlic
    4 tbsp olive oil
    1 tsp lemon juice
    2 red bell peppers"
  )) %>%
  #Format
  separate(., tmp, into = c("recipe_name", "Ingredients"), sep = ";") %>%
  separate_rows(., Ingredients, sep = "\\n") %>%
  #Standardise food and unit names and amounts, turn volume to weight and calculate nutrient content
  standardiseFoodList() %>%
  findFoodInDatabase(., "volume_weight") %>%
  calculateWeightOfIngredients() %>%
  findFoodInDatabase(., "nutrients") %>%
  #Calculate nutrients per 100 g as this will be used as additional entries to calculate nutrient content of other recipes
  calculateNutrientContentOfFoodlist(., calculate_nutrients = "per 100 g") %>%
  #Format as new database entries for the nutrients database
  createNewDatabaseEntry(., "nutrients")
```
The output from calculateWeightOfIngredients is used to calculate total nutrient content (default of calculateNutrientConcentOfFoodlist) of the recipes:
```{r nutrient-environmental-calculations, message=FALSE, results = FALSE}
nutrient_content <- findFoodInDatabase(weighted_ingredients, "nutrients", additional_entries = nutrients_additional_ingredients) %>%
  calculateNutrientContentOfFoodlist()
```


